May I Ask Where You Have Been, Photon?

### 2021-11-19 article.tex
add double slit figures, almost finish mzi explanation

@@

Quantum physics describes physical interactions that happen at the atomic and
sub-atomic levels, and underpins the foundation for most of modern technology.
However, physicists are still unsure how to use standard quantum theory to 
answer the following question: what path did a photon take through an 
interferometer before quantum measurement was performed?

@@

Another apparatus that can demonstrate single-particle interference is the
Mach-Zehnder interferometer, which consists of a light source and a detector 
along with two beam splitters (basically half-silvered mirrors) and two 
mirrors, arranged as shown in [figure]. In the classical version of the 
experiment, the light source emits a continuous beam of light, which partly 
passes through the the first beam splitter, and is then reflected by the two 
mirrors to meet at the second beam splitter. Reflection at an interface between 
vacuum and the silvered side of a mirror causes a phase change through a purely 
classical phenomenon, which means that the beams coming out of the second beam 
splitter combine totally constructively on one side, and totally destructively 
on the other, in analogy to the bright and dark bands in the classical double 
slit experiment. Extending the analogy to the double-slit experiment, if the 
light source in the Mach-Zehnder interferometer is tuned to a low-enough 
intensity that single photons emitted can be resolved by the detector, it is 
found that the photon arrives at the detector 100\% of the time, demonstrating 
the phenomenon of a particle interfering with itself. Here, if we ask the 
photon where it has been, our "common sense" approach to quantum mechanics says 
that it has been in both the upper and the lower arm of the interferometer.

>>>>

\usepackage{svg}
\usepackage{subfig}

\newcommand{\Q}{\medskip\bfseries Q: }
\newcommand{\A}{\par\textbf{A:} \normalfont}

@@

\begin{figure}[ht]
	\includesvg[width=\linewidth]{figures/interferometer1.svg}
	\caption{Simple Mach Zehnder interferometer}
	\label{fig:interferometer1}
\end{figure}

Quantum physics describes physical interactions that happen at the atomic and
sub-atomic levels, and underpins the foundation for most of modern technology.
However, physicists are still unsure how to use standard quantum theory to 
answer the following question: what path did a photon take through an 
interferometer before quantum measurement was performed?  A recent experiment 
conducted by A. Danan and colleagues at Tel Aviv University in Israel sheds 
light on how difficult it is exactly to answer this question in any but the 
simplest of situations.

@@

In the double slit experiment (as performed by Thomas Young), a beam of light 
is passed through two closely spaced slits and projected onto the screen.  
According to classical wave theory, the intensity of light detected at the 
screen is given by the intensity of the incident waves as well as their 
relative phases. Because of the different different distances traveled by 
different wavefronts, phase difference arise, in turn giving rise to an 
interference pattern on the screen, as shown in figure 
\label{fig:uncollimated}.

\begin{figure}
	\fbox{\subfloat[Standard 
	setup]{\includesvg[width=0.5\linewidth]{figures/double-slit.svg}}\label{fig:uncollimated}}
	\fbox{\subfloat[Collimated 
	beams]{\includesvg[width=0.5\linewidth]{figures/collimated-double-slit.svg}}\label{fig:collimated}}
	\caption{Young's double slit interference experiment}
	\label{fig:double-slit}
\end{figure}

Another apparatus that can demonstrate single-particle interference is the
Mach-Zehnder interferometer, which consists of a light source and a detector 
along with two beam splitters (basically half-silvered mirrors) and two 
mirrors, arranged as shown in [figure]. In the classical version of the 
experiment, the light source emits a continuous beam of light, which partly 
passes through the the first beam splitter, and is then reflected by the two 
mirrors to meet at the second beam splitter. Reflection at an interface between 
vacuum and the silvered side of a mirror causes a phase change through a purely 
classical phenomenon, which means that the beams coming out of the second beam 
splitter combine totally constructively on one side, and totally destructively 
on the other, in analogy to the bright and dark bands in the classical double 
slit experiment. Extending the analogy to the double-slit experiment, if the 
light source in the Mach-Zehnder interferometer is tuned to a low-enough 
intensity that single photons emitted can be resolved by the detector, it is 
found that the photon arrives at the detector 100\% of the time, demonstrating 
the phenomenon of a particle interfering with itself. Of course, if we remove 
the second beam splitter, there is a 50\% chance that the photon will show up 
in our detector, and a 50\% chance that it will come out of the other branch of 
the interferometer.

Here, we attempt to answer the titular question using the "common sense" 
approach amateur science enthusiasts and undergraduate physics students are 
familiar with:

\Q May I Ask Where You Have Been, Photon?

\A The photon's trajectory in the interferometer is a superposition of the 
classical trajectories through the upper and lower arms of the 
interferometer.\medskip

The Tel-Aviv experiment starts off with a modified version of the Mach-Zehnder 
interferometer, in which each mirror oscillates vertically with a unique 
frequency.  Additionally, instead of a simple photodetector, a stacked array of 
detectors is used, making it possible to monitor the frequency spectrum of the 
detected light wave. As expected, the power spectrum shows equal contributions 
at frequencies $f_A$ and $f_B$ associated with mirrors $A$ and $B$.

Whilst this does not directly answer our question, since the photon source 
emits numerous photons, which means that different ones could have caused the 
peaks at $f_A$ and $f_B$, it at least confirms that some of the photons have 
been at each mirror.

In the second variation of the experiment, the same steps are repeated, but 
with the second beam splitter removed. This time, only the peak at $f_B$ 
remains, again confirming our interpretation of the Mach-Zehnder interferometer 
experiment.

Next, a larger interferometer is built, as illustrated in figure 
\ref{fig:danan-2a}. There is a nested Mach-Zehnder interferometer in the upper 
arm, and the beam intensity is now split 1:2 between the upper and the lower 
arm, so the ratio of intensities at the vibrating mirrors $A$, $B$, and $C$ is 
1:1:1.  Again, the power spectrum shows what we would expect: peaks at all 
frequencies, with larger peaks at $f_E$ and $f_F$.

So far, the results of this experiment have confirmed our intuitive, "common 
sense" approach to accounting for a quantum particle's trajectory as a sum of 
all possible paths. The surprising result from this experiment is obtained when 
the following changes are made, as illustrated in figure \ref{fig:danan-2b}:
either the second beam splitter of the mini-Mach Zehnder interferometer is 
flipped, or mirror $B$ is moved to slightly adjust the path length, causing 
destructive interference of the light propagating towards $F$.

\begin{figure}
	\subfloat[]{\includesvg[width=0.5\linewidth]{figures/danan-2a.svg}\label{fig:danan-2a}}
	\subfloat[]{\includesvg[width=0.5\linewidth]{figures/danan-2b.svg}\label{fig:danan-2b}}
	\newline
	\hspace*{\fill}
	\subfloat[]{\includesvg[width=0.5\linewidth]{figures/danan-2c.svg}\label{fig:danan-2c}}
	\hspace*{\fill}
	\caption{Nested Mach-Zehnder interferometer}
	\label{fig:danan}
\end{figure}
